### 2025-06-19 lecturenotes_aima104/ch3_aima104.tex
ch3aima104

\subsection{Exercise}
Diagonalizable the following matrix, if possible.
$$A=
\begin{bmatrix}
  5 & 0 & 0 & 0 \\
  0 & 5 & 0 & 0 \\
  1 & 4 & -3 & 0 \\
  -1 & -2 & 0 & -3
\end{bmatrix}
$$

\subsection{Similarity}
If $A$ an $B$ are $n \times n$ matrices, then $A$ \textbf{similar to} $B$ if there is an invertible matrix $P$ such that $$A = PBP^{-1}$$ or equivalently $B=P^{-1}AP$. Writing $Q=P^{-1}$, we can say that $B$ is similar to $A$, and we simply say $A$ and $B$ are similar.
\begin{theorem}
  If $n \times n$ matrices $A$ and $B$ are similar, then they have the same characteristic polynomial and hence the same eigenvalues (with the same multiplicities).
\end{theorem}
\begin{proof}
 \hspace{1cm} Given, $B=P^{-1}AP$ we have, $(B-\lambda I) = P^{-1}AP - \lambda P^{-1}P = P^{-1}(A- \lambda I) P$. Then
    \begin{align*}
      det(B-\lambda I) &= \det[P^{-1}(A- \lambda I) P] \\
                       &= \det(P^{-1}).\, \det(A- \lambda I).\, \det(P) \\
                       &= \det(P^{-1}).\, \det(P). \, \det(A- \lambda I) \\
                       &= \det(A- \lambda I)
    \end{align*}
  As, $\det(P^{-1}).\, \det(P) = \det(P^{-1}.P)=det(I) = 1$. Hence $A$ and $B$ has same characteristic polynomial. And the proof follows.
\end{proof}
\end{document}

>>>>

\subsection{Multiplicities of an eigenvalue}
\begin{itemize}
\item The algebraic multiplicity of $\lambda$ is the number of times $(\lambda - t)$ occurs as a factor of in its characteristic polynomial.
\item The geometric multiplicity of an eigenvalue $\lambda$ is the dimension of the corresponding eigenspace $E_{\lambda}$. i.e $dim\{E_{\lambda}\}$.
\end{itemize}
\subsection{Exercise}
Diagonalizable the following matrix, if possible.
$$A=
\begin{bmatrix}
  4 & 0 & 0  \\
  1 & 4 & 0  \\
   0& 0 & 5  \\
\end{bmatrix}
$$

\subsection{Similarity}
If $A$ an $B$ are $n \times n$ matrices, then $A$ \textbf{similar to} $B$ if there is an invertible matrix $P$ such that $$A = PBP^{-1}$$ or equivalently $B=P^{-1}AP$. Writing $Q=P^{-1}$, we can say that $B$ is similar to $A$, and we simply say $A$ and $B$ are similar.
\begin{theorem}
  If $n \times n$ matrices $A$ and $B$ are similar, then they have the same characteristic polynomial and hence the same eigenvalues (with the same multiplicities).
\end{theorem}
\begin{proof}
 \hspace{1cm} Given, $B=P^{-1}AP$ we have, $(B-\lambda I) = P^{-1}AP - \lambda P^{-1}P = P^{-1}(A- \lambda I) P$. Then
    \begin{align*}
      det(B-\lambda I) &= \det[P^{-1}(A- \lambda I) P] \\
                       &= \det(P^{-1}).\, \det(A- \lambda I).\, \det(P) \\
                       &= \det(P^{-1}).\, \det(P). \, \det(A- \lambda I) \\
                       &= \det(A- \lambda I)
    \end{align*}
  As, $\det(P^{-1}).\, \det(P) = \det(P^{-1}.P)=det(I) = 1$. Hence $A$ and $B$ has same characteristic polynomial. And the proof follows.
\end{proof}

\section{Linear Transformation and Diagonalization}
We will see that the transformation $x \mapsto As$ is essentially same as the very simple mapping $u \maps to Du$ when viewed from the proper perspective.
\begin{theorem}
  Suppose $A = PDP^{-1}$, where $D$ is a diagonal $n \times n$. If $B$ is the basis for $\mathbb{R}$, formed from the columns of $P$, then $D$ is the $B$-matix for the transformation $x \mapsto Ax$.
\end{theorem}
\end{document}

### 2025-06-19 lecturenotes_aima203/ch4_2_aima203.tex
ch3aima104

\subsection{Linear Splines}
Suppose the given data points are $(x_i,y_i), \; i=0,1,\, ..., \, n$ and  $h_i = x_i-x_{i-1} , \; i=1,\, 2, \, ..., \, n$. \\
Let $s_i(x)$ be a straight line from $x_{i-1}$ to $x_i$. Then, the slope of $s_i(x)$ is \\
$\displaystyle m_i = \frac{y_i - y_{i-1}}{x_i - x_{i-1}}$ and $s_i(x)= y_i + m_i(x-x_i)$. \\[2mm]
From the discussion above the $s_i(x)$ are the \textbf{linear splines}.

@@

\subsection{Quadratic Splines}
Let $s_i(x)$ be a quadratic approximation of the data points in the sub-interval $[x_{i-1} -x_i]$ satisfying the following conditions:
\begin{enumerate}
\item $s_i(x)$ and $s_i'(x)$ are continuous on $[x_0,x_n]$,
\item $s_i(x)=y_i, \hspace{1cm} i=0,\, 1,\, 2, \, ..., \, n$
\item $s_i'(x) = \frac{1}{h_i} [ (x_i -x)m_{i-1} + (x-x_{i-1})m_i$ as $s_i'(x)$ is linear, where $m_i = s_i'(x)$.
\end{enumerate}
\textbf{Integrating} $s_i'(x)$ with respect to $x$, we obtain
\begin{equation}
  s_i(x) = \frac{1}{h_i} \left [ -\frac{(x_i -x)^2}{2}\, m_{i-1} + \frac{(x-x_{i-1})^2}{2} \, m_i \right ] + c_i
\end{equation}
Putting $x=x_{i-1}$ we get,
\begin{equation}
  c_i = y_{i-1} + \frac{h_i}{2}m_{i-1}
\end{equation}
Imposing the continuity condition on the spline functions $s_i(x)$ and imposing the natural spline condition $s''_1(x_1)=0$ we obtain the following realtions to calculate $m_i$.
\begin{equation}
  m_{i-1} +m_{i} = \frac{2}{h_i}(y_i -y_{i-1}), \hspace{1cm} i = 1,\, 2, \, ..., \, n \hspace{3mm} \text{and} \hspace{3mm} m_0 = m_1
\end{equation}

>>>>

\subsection{Linear Splines}
Suppose the given data points are $(x_i,y_i), \; i=0,1,\, ..., \, n$ and  $h_i = x_i-x_{i-1} , \; i=1,\, 2, \, ..., \, n$. \\
Let $s_i(x)$ be a straight line from $x_{i-1}$ to $x_i$. Then, the slope of $s_i(x)$ is \\
$\displaystyle m_i = \frac{y_i - y_{i-1}}{x_i - x_{i-1}}$ and $s_i(x)= y_{i-1} + m_i(x-x_{i-1})$. \\[2mm]
From the discussion above the $s_i(x)$ are the \textbf{linear splines}.

@@

\subsection{Quadratic Splines}
Let $s_i(x)$ be a quadratic approximation of the data points in the sub-interval $[x_{i-1} -x_i]$ satisfying the following conditions:
\begin{enumerate}
\item $s_i(x)$ and $s_i'(x)$ are continuous on $[x_0,x_n]$,
\item $s_i(x_i)=y_i, \hspace{1cm} i=0,\, 1,\, 2, \, ..., \, n$
\item $s_i'(x) = \frac{1}{h_i} [ (x_i -x)m_{i-1} + (x-x_{i-1})m_i ]$ as $s_i'(x)$ is linear, where $m_i = s_i'(x)$.
\end{enumerate}
\textbf{Integrating} $s_i'(x)$ with respect to $x$, we obtain
\begin{equation}
  s_i(x) = \frac{1}{h_i} \left [ -\frac{(x_i -x)^2}{2}\, m_{i-1} + \frac{(x-x_{i-1})^2}{2} \, m_i \right ] + c_i
\end{equation}
Putting $x=x_{i-1}$ we get,
\begin{equation}
  c_i = y_{i-1} + \frac{h_i}{2}m_{i-1}
\end{equation}
Imposing the continuity condition on the spline functions $s_i(x)$ we get,
\begin{equation}
  m_{i-1} +m_{i} = \frac{2}{h_i}(y_i -y_{i-1}), \hspace{2cm} i = 1,\, 2, \, ..., \, n
\end{equation}
Imposing the natural spline condition $s''_1(x_1)=0$ we obtain, $ m_0 = m_1$
